The Early Eocene Climate Optimum reconstructed with Bayesian multi-proxy integration

### 2022-09-03 manuscript/manuscript.Rmd
manuscript methods: priors. small update to gradient: sdy = 0 (default)

This structure implies that $\mu_j$ is not fixed at the mean of the temperatures $t_{i=1,...,m,~j}$ of location $j$, but is drawn (`regularised`) towards the overall logistic regression curve. The pull towards the regression curve tends to be high when $m$ is low, when the observations $t_{i=1,...,m,~j}$ are scattered, and/or when the overall standard deviation $\sigma$ is low. In practice, this has the desirable consequence that locations with few observations and large temperature differences between observations have less influence on the regression curve than well-sampled locations with consistent temperature observations.

**Priors.** In the Bayesian framework, priors need to be placed on the unknown parameters of a model. For $\sigma$ and $\sigma_{j=1,..n}$ we use weakly informative inverse-gamma priors with shape parameter $alpha$ = ... and scale parameter $beta$ = ... . Moderately changing these priors has little effect on the model outcome, which is desirable as we have little a-priori knowledge of these parameters. In contrast, we put informative priors on the regression coefficients $A$, $K$, $M$ and $B$, based on basic physical principles. Specifically, we put a skew-normal prior on the lower asymptote $A$, with ...  . This constrains $A$ to not be $<< -2^\circ C$, which is the freezing point of sea water. We chose this distribution such that its mode is $~ 0^\circ C$, slightly higher than the modern...

>>>>

This structure implies that $\mu_j$ is not fixed at the mean of the temperatures $t_{i=1,...,m,~j}$ of location $j$, but is drawn (`regularised`) towards the overall logistic regression curve. The pull towards the regression curve tends to be high when $m$ is low, when the observations $t_{i=1,...,m,~j}$ are scattered, and/or when the overall standard deviation $\sigma$ is low. In practice, this has the desirable consequence that locations with few observations and large temperature differences between observations have less influence on the overall regression than well-sampled locations with consistent temperature observations.

**Priors.** In the Bayesian framework, priors need to be placed on the unknown parameters of a model. We placed weakly informative, conjugate inverse-gamma priors on $\sigma$ and $\sigma_{j=1,..n}$:
$$\sigma  \sim \sqrt{Inv{\text -}Gamma(\alpha+\frac{n}{2},\beta + 0.5\times(\mu_j-\nu_j))}, ~~~~~ j = 1,...,n,  $$
$$\sigma_j  \sim \sqrt{Inv{\text -}Gamma(\alpha+\frac{m}{2},\beta + 0.5\times(t_{i,j}-\mu_j))}, ~~~~~ i = 1,...,m,~~~~~j = 1,...,n. $$
We set $\alpha = 1$ and $\beta = 1$, allowing these priors to be quickly overwhelmed by the data as $n$ and $m$ increase, which is desirable as we have little a-priori knowledge of these parameters.

In contrast, we put informative priors on the regression coefficients $A$, $K$, $M$ and $B$, based on physical principles and knowledge from the modern climate system. Specifically, we used the priors to express the following constraints and prior knowledge:
1) Predicted seawater surface temperatures is not allowed to be $<< -2^\circ C$, the freezing point of sea water.
2) Input of solar energy decreases from the tropics to the poles. Hence, the latitudinal temperature gradient is generally negative, i.e. temperature decreases with absolute latitude. This is achieved by constraining $K-A$ to be positive.
3) The steepness of the gradient is likely highest in mid-latitudes.
4) The 
a skew-normal prior on the lower asymptote $A$, with ...  . This constrains $A$ to not be $<< -2^\circ C$, which is the freezing point of sea water. We chose this distribution such that its mode is $~ 0^\circ C$, slightly higher than the modern...
